---
title: "Assignment4"
author: "Kexin Wu"
date: "04/12/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Assignment 4 for Statistical Computing and Empirical Mathods

### 1 Bayes theorem

 Let $A$ be the event that it rains next week and $B$ the event that the weather
forecaster predicts that there will be rain next week. Let’s suppose that the 
probability of rain next week is  $\mathbb{P}(A) = 0.9$. Suppose also that the 
conditional probability that there is a forecast of rain, given that it really does 
rain, is $\mathbb{P}(B\mid{A}) = 0.8$. On the other hand, the conditional 
probability that there is a forecast of dry weather, given that there really 
isn’t any rain is $\mathbb{P}(B^c \mid {A^c}) = 0.75$.
 
Now suppose that there is a forecast of rain. What is the conditional probability 
of rain, given the forecast of rain $\mathbb{P}(A\mid{B})$?

Answer:

First, we compute $\mathbb{P}(B)$:


$$\mathbb{P}(B) = \mathbb{P}({B}\cap{A}) + \mathbb{P}({B}\cap{A^c})
     = \mathbb{P}({B}\mid{A})\mathbb{P}(A) + \mathbb{P}({B}\mid{A^c})\mathbb{P}(A^c)
     = 0.8 * 0.9 + {1 - \mathbb{P}({B^c}\mid{A^c})}\mathbb{P}(A^c)
     = 0.72 + 0.025 = 0.745$$
     
We then apply Bayes theorem:

$$\mathbb{P}({A}\mid{B}) = \frac{\mathbb{P}(A)\mathbb{P}({B}\mid{A})}{\mathbb{P}(B}) = \frac{0.9 * 0.8}{0.745} \approx 0.966$$

       
### 2 Conditional probabilities

Suppose we have a probability space ($\Omega$, $\varepsilon$, $\mathbb{P}$).

*(Q1)* Suppose that $A$, $B$ $\in$ $\varepsilon$ and $A \subseteq B$. Give an 
expression for $\mathbb{P}({A}\mid{B})$ in terms of $P(A)$ and $P(B)$. What
about when $\mathbb{P}(A \backslash B) = 0$?

*(A1)* If 
$$\mathbb{P}(A \backslash B) = 0$$
Then,
 $$\mathbb{P}(A) = \mathbb{P}(A \cap {B}) + \mathbb{P}(A \cap {B^c})
 = \mathbb{P}(A \cap {B}) + \mathbb{P}(A \backslash {B})
 = \mathbb{P}(A \cap {B})$$
 
 Hence, 
 $$\mathbb{P}(A \mid B) = \frac{\mathbb{P}(A \cap {B})}{\mathbb{P}(B)} = \frac{\mathbb{P}(A)}{\mathbb{P}(B)}$$.

------

*(Q2)* Suppose that $A$, $B$ $\in$ $\varepsilon$ with $A \cap B=\phi$. Give an 
expression for $\mathbb{P}({A}\mid{B})$. What about when $\mathbb{P}(A\cap{B}) = 0$? 

*(A2)*  If 
$$\mathbb{P}(A \cap {B}) = 0$$
Then,
$$\mathbb{P}(A \mid {B}) = \frac{\mathbb{P}(A \cap {B})}{\mathbb{P}(B) }= 0$$

------

*(Q3)* Suppose that $A$, $B$ $\in$ $\varepsilon$ with $B \subseteq A$. Give an 
expression for $\mathbb{P}(A\mid{B})$. What about when $\mathbb{P}(B \backslash A) = 0$?

*(A3)*  If 
$$\mathbb{P}(B \backslash A) = 0$$
Then,
$$\mathbb{P}(A^c \cap {B}) = 0$$
 
Hence, 
$$\mathbb{P}(B) = \mathbb{P}(A \cap {B}) + \mathbb{P}(A^c \cap {B}) = \mathbb{P}(A \cap {B})$$
And,
$$\mathbb{P}(A \mid B) = \frac{\mathbb{P}(A \cap {B})}{\mathbb{P}(B)} = \frac{\mathbb{P}(B)}{\mathbb{P}(B)} = 1$$

------

*(Q4)* Suppose that $A \in \varepsilon$. Give an expression for $\mathbb{P}(A \mid \Omega)$ in terms of $\mathbb{P}(A)$.

*(A4)* In this case, we have $A \subseteq \Omega$ and $\mathbb{P}(\Omega) = 1$, 
So,
$$\mathbb{P}(A \mid \Omega) = \frac{\mathbb{P}(A)}{\mathbb{P}(\Omega)} = \mathbb{P}(A)$$

------

*(Q5)* Show that given three events $A$, $B$, $C$ $\in \varepsilon$, we have $\mathbb{P}(A \bigcap {B} \bigcap {C}) = \mathbb{P}(A \mid {B} \bigcap {C}) \cdot \mathbb{P}(B \mid {C}) \cdot \mathbb{P}(C)$.

*(A5)*  $$\mathbb{P}(A \cap {B} \cap {C}) = \mathbb{P}(A \cap (B \cap C)
= \mathbb{P}(A \mid (B \cap C)) \cdot \mathbb{P}(B \cap C)
= \mathbb{P}(A \mid (B \cap C)) \cdot \mathbb{P}(B \mid C) \cdot \mathbb{P}(C)$$

------

*(Q6)* Show that given three events $A$, $B$, $C \in \varepsilon$ we have $\mathbb{P}(A \mid B \cap C) = \frac{\mathbb{P}(B \mid A \cap C) \cdot \mathbb{P}(A \mid C)}{\mathbb{P}(B \mid C)}$.

*(A6)*  First, we compute the left part of the formula that 
$$\mathbb{P}(A \mid B \cap C) = \frac{\mathbb{P}(A \cap B \cap C)}{\mathbb{P}(B \cap C)}$$

   Then, we compute the right part, since that 
   $$\mathbb{P}(A \mid C) = \frac{\mathbb{P}(A \cap C)}{\mathbb{P}(C)}$$
   And,
   $$\mathbb{P}(B \mid C) = \frac{\mathbb{P}(B \cap C)}{\mathbb{P}(C)}$$
   
   So,
   $$\frac{\mathbb{P}(B \mid A \cap C) \cdot \mathbb{P}(A \mid C)}{ \mathbb{P}(B \mid C)}
   = \frac{\mathbb{P}(B \mid A \cap C) \cdot \mathbb{P}(A \cap C)}{\mathbb{P}(B \cap C)}$$
   
   Hence,
   $$\mathbb{P}(B \mid A \cap C) \cdot \mathbb{P}(A \cap C) = \mathbb{P}(A \cap B \cap C)$$
   
   $$\frac{\mathbb{P}(B \mid A \cap C) \cdot \mathbb{P}(A \mid C)}{\mathbb{P}(B \mid C)}
   = \frac{\mathbb{P}(A \cap B \cap C)}{\mathbb{P}(B \cap C)}
   = \mathbb{P}(A \mid B \cap C)$$
   

### 3 Sampling with replacement
 Recall that for positive integers $n$ and $k$, $\binom{n}{k} = \frac{n!}{(k!)(n-k)!}$ 
 gives the number of subsets of size $k$ from $a$ set of $n$ objects.

 You can compute this number straightforwardly within R via the choose function choose( ). For example,
if we want to compute the number of different subsets of size 3 from a collection of size 8 we would compute:
choose(8,3)
```{r}
choose(8, 3)
```

Suppose we have a bag containing 10 spheres. This includes 3 red spheres and 7 blue spheres.

Let’s suppose that we draw a sphere at random from the bag (all spheres have equal probability of being drawn).
We record its colour and then return the sphere to the bag. This process is repeated 35 times. 
This is an example of sampling with replacement since the spheres are replaced after each draw.

*(Q)* Write down a mathematical expression for the probability that $z$ out of 
the 35 selections were red spheres (here $z \in \{1,...,35\}$).
Try doing this with in laTex, making use of the laTex functions binom{}{} and frac{}{}.

*(A)* The expression is:
$$\binom{35}{z} \cdot (\frac{3}{10}) ^ z \cdot (\frac{7}{10}) ^{35-z}$$

*(Q)* Next write an R function called prob_red_spheres() which takes $z$ as an 
argument and computes the probability that $z$ out of a total of the 35 balls 
selected are red.

*(A)* 
```{r}
num_red_balls <- 3
num_total_balls <- 10
total_draws <- 35

prob_red_spheres <- function(z){
  result <- choose(total_draws, z) * 
    (num_red_balls / num_total_balls) ^ z * 
    ((num_total_balls - num_red_balls) / num_total_balls) ^ (total_draws - z)
  return(result)
}
```

Test my function as follows:
```{r}
prob_red_spheres(20)
```

*(Q)* Generate a data frame called prob_by_num_reds with two columns num_reds and prob. 
The num_reds column should contain numbers 1 through 35 and the prob column should give 
the associated probability of selecting that many reds out of a total number of 35 selections.

*(A)*
```{r}
library(tidyverse)
```

```{r}
prob_by_num_reds <- data.frame(num_reds = seq(35)) %>%
  mutate(prob = prob_red_spheres(num_reds))
```

Display the first 3 rows of your data frame:
```{r}
prob_by_num_reds %>% head(3)
```

*(Q)* Now use the geom_line() function within the ggplot2 library, in conjunction 
with your data frame to display a plot of the probability as a function of the
number of reds.

*(A)*
```{r}
prob_by_num_reds %>% 
  ggplot(mapping = aes(x = num_reds, y = prob)) + 
  geom_line() +
  theme_bw() +
  xlab("Number of reds") + 
  ylab("Probability")
```

*(Q)* Next we shall explore the sample() function within R. Let’s suppose we want
to simulate a random experi- ment in which we sample with replacement from a 
collection of 10 objects, and repeat this process 35 times. We can do this by calling:

```{r}
sample(10 , 35, replace=TRUE)
```

Try this out for yourself. The output should be a vector of length 35 consisting 
entirely of numbers between 1 and 10. Since this is sampling with replacements 
and the number of samples exceeds the number of elements there will be repetitions.

Try rerunning the function. You probably get a different sample. This is to be 
expected, and even desirable, since the function simulates a random sample. 
However, the fact that we get a different answer every time we run the code is 
problematic from the prespective of reproducibility. To avoid this process we 
can set a random seed via the function set.seed(). By doing so we should get the 
same output every time. Try the following out for your self:
```{r}
for (i in 1:5) {
  print(sample(100, 5, replace = FALSE))
  # The result may well differ every time
}

## Resetting the random seed every time
for(i in 1:5){
  set.seed(1) 
  print(sample(100,5,replace=FALSE)) # The result should not change
}
```

```{r}
set.seed(0)
for (i in 1:5) {
  print(sample(100, 5, replace = FALSE))
  # The result may well differ every time
}
```

We shall now use the sample() to construct a simulation study to explore the 
probability of selecting z red balls from a bag of size 10, with 3 red and 7 
blue balls, when sampling 35 balls with replacement.

First set a random seed. Then create a data frame called sampling_with_replacement_simulation 
consisting of a two columns. The first is called trial and contains numbers 1 
through 1000. The second is called sample_balls and corresponds to a random 
sample of size 35 from a bag of size 10, with replacement. We can do this as follows:
```{r}
num_trials <- 1000 # set the number of trials
set.seed(0) # set the random seed
sampling_with_replacement_simulation <- data.frame(trial = 1:num_trials) %>%
  mutate(sample_balls = map(.x = trial, ~sample(10, 35, replace = TRUE)))
# generate collection of num_trials simulations
```

Now add a new column called num_reds such that, for each row, num_reds contains 
an integer which gives the number of items within the sample for that row (the
entry in the sample_balls column) which are less than or equal to three. For 
example, suppose that some row of the data frame, the sample_balls column contains 
the following list:

9, 4, 7, 1, 2, 7, 2, 3, 1, 5, 5, 10, 6, 10, 7, 9, 5, 5, 9, 9,
5, 5, 2, 10, 9, 1, 4, 3, 6, 10, 10, 6, 4, 4, 10

Then the corresponding row of the num_reds column should contain the number 8, 
since 8 of these values are less than equal to 3. You may want to use the functions 
mutate(), map_dbl and sum().

```{r}
sampling_with_replacement_simulation <- sampling_with_replacement_simulation %>%
  mutate(num_reds = map_dbl(.x = sample_balls, ~sum(.x <= 3)))
```

### Help!!

Next we shall add a new column called simulation_count to our existing data frame 
prob_by_num_reds which gives the number of times within our simulation we observed 
the corresponding number of reds. We can do this as follows:
```{r}
num_reds_in_simulation <- sampling_with_replacement_simulation %>% pull(num_reds)
# we extract a vector corresponding to the number of reds in each trial
prob_by_num_reds <- prob_by_num_reds %>%
  mutate(simulation_count = map_dbl(.x = num_reds, ~sum(num_reds_in_simulation == .x)))
# add a column which gives the number of trials with a given number of reds
prob_by_num_reds %>% head(5)
```

Next we add a column called expected_count corresponding to the expected number 
of observed reds in 1000 based upon your probability formula.
```{r}
prob_by_num_reds <- prob_by_num_reds %>%
  mutate(expected_count = num_trials * prob)
prob_by_num_reds %>% head(5)
```

Finally, create a plot which compares the results of your simulation with the 
expected count based on your probability formula. The concept of expectation will
be discussed in Lecture 11.

Your result should look something like the plot below. Of course, since this is a
random simulation, your result may well look slightly different.
```{r}
prob_by_num_reds %>%
  rename(Simulation = simulation_count, Expected = expected_count) %>%
  pivot_longer(cols = c("Simulation", "Expected"),
               names_to = "Type", values_to = "count") %>%
  ggplot(aes(num_reds, count)) + 
  geom_line(aes(linetype = Type, color = Type)) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  theme_bw() +
  xlab("Number of reds") +
  ylab("Count")
```

### 4 Sampling without replacement

This question is more challenging. However, you should aim to complete at least 
the simulation component using ideas from the previous question.

Let’s suppose we have large bag containing 100 spheres. There are 50 red spheres, 
30 blue spheres and 20 green spheres. Suppose that we sample 10 spheres from the 
bag without replacement.

What is the probability that one or more colours is missing from your selection?

First aim to answer this question via a simulation study using ideas from the previous question.

You may want to use the following steps:

1. First set a random seed;
2. Next set a number of trials, a number of reds, a number of blues, a number of 
greens and a sample size;
3. Now use a combination of the functions sample(), mutate() and map() to generate 
your samples.Here you are creating sample of size 10 from a collection of 100 balls -
the sampling is done without replacement;
4. Now compute the number of “reds”, “greens” and “blues” in your sample using 
the map_dbl() and mutate() functions;
5. Compute the minimum of the three counts using the pmin() function. When this 
minimum is zero, then one of the three colours is missing. It is recommended that 
you look up the difference between pmin() and min() here;
6. Compute the proportion of rows for which the minimum number of the three counts is zero.

*(Q)*
Try this initially with a small number of simulations. Increase your number of 
simulations to about a relatively large number to get a more accurate answer,
once everything seems to be working well.

*(A)*
```{r}
set.seed(0) # set the random seed
num_trials <- 100000 # se0t the number of trials
num_red <- 50 # set the number of reds
num_blue <- 30 # set the number of blues
num_green <- 20 # set the number of greens
sample_size <- 10

total <- num_red + num_blue + num_green

sampling_without_replacement_simulation <- data.frame(trial = 1:num_trials) %>%
  mutate(sample_balls = map(.x = trial, ~sample(total, sample_size, replace = FALSE))) %>%
  # simulate without replacement the sampling of 10 balls from a bag of 100
  mutate(num_reds = map_dbl(.x = sample_balls, ~sum(.x <= num_red))) %>%
  mutate(num_blues = map_dbl(.x = sample_balls, ~sum((.x > num_red)  & (.x <= num_red + num_blue)))) %>%
  mutate(num_greens = map_dbl(.x = sample_balls, ~sum((.x > num_red + num_blue) & (.x <= total)))) %>%
  # count the number of each colour
  mutate(colour_missing = (pmin(num_reds, num_blues, num_greens) == 0))
  # look for missing colours

prob_missing_simulation <- mean(sampling_without_replacement_simulation %>%
                                  pull(colour_missing))

```

*(Q)* 
The next part of the question is more challenging and may be omitted if you
are short on time.
(*) Once you have a simulation based answer you can try and use “combinations” 
with $\binom{n}{k}$ to compute the k probability directly. First aim and compute 
the number of subsets of size 10 from 100 which either entirely miss out one of 
the subsets Reds = {1,...,50}, Blues = {51,...,80}, Greens = {81,...,100}. Be careful 
not to double count some of these subsets! Once you have computed all such subsets 
combine with the formula for the total number of subsets of size 10 from a set of 100, 
to compute the probability of missing a colour.

*(A)*
 We first compute the number of subsets of size 10 containing no reds, no blues and no greens:
 
 1. N~R~ - The number of subsets with no reds is $\binom{50}{10}$
 2. N~B~ - The number of subsets with no blues is $\binom{70}{10}$
 3. N~G~ - The number of subsets with no greens is $\binom{80}{10}$
 
 Next we count the number of all red, all blue, and all green:
 
 1. M~R~ - The number of subsets with all reds is $\binom{50}{10}$
 2. M~B~ - The number of subsets with all blues is $\binom{30}{10}$
 3. M~G~ - The number of subsets with all greens is $\binom{20}{10}$
 
 Now the total number of subsets with a missing colour is
 $$Q = (N_R + N_B + N_G) - (M_R + M_B + M_G)$$
 The total of subsets of size 10 from a collection of a 100 is $T = \binom{100}{10}$.
 
 Dividing by the total number of subsets gives us the probability $Q / T \approx 0.118$

```{r}
NR = choose(total - num_red, sample_size)
NB = choose(total - num_blue, sample_size)
NG = choose(total - num_green, sample_size)

MR = choose(num_red, sample_size)
MB = choose(num_blue, sample_size)
MG = choose(num_green, sample_size)

Q = (NR + NB + NG) - (MR + MB + MG)
TS = choose(total, sample_size)

prob_missing_theory = Q / TS

prob_missing_simulation

prob_missing_theory
```

### 5 Mutual independence and pair-wise independent

Consider a simple probability space $(\Omega, \varepsilon, \mathbb{P})$ with $\Omega = \{(0, 0, 0), (0, 1, 1), (1, 0, 1), (1, 1, 0)\}$.

Since $(\Omega, \varepsilon, \mathbb{P})$ is a simple probability space containing four elements we have

$$ \mathbb{P}(\{(0, 0, 0\}) = \mathbb{P}(\{(0, 1, 1\}) = \mathbb{P}(\{(1, 0, 1\}) = \mathbb{P}(\{(1, 1, 0\}) = \frac{1}{4}$$
Consider the events $A := \{(1, 0, 1), (1, 1, 0)\}$, $B :=  \{(0, 1, 1), (1, 1, 0)\}$ and $C := \{(0, 1, 1), (1, 0, 1)\}$.

*(Q)* Verify that $\mathbb{P}(A \cap B) = \mathbb{P}(A) \cdot \mathbb{P}(B)$, 
$\mathbb{P}(A \cap C) = \mathbb{P}(A) \cdot \mathbb{P}(C)$ and  
$\mathbb{P}(B \cap C) = \mathbb{P}(B) \cdot \mathbb{P}(C)$. 
Hence, we deduce that the events $A$, $B$, $C$ are pair-wise indepenent.

*(A)* Since, $$A \cap B = \{(1, 0, 1\}, A \cap C = \{(1, 0, 1\}, B \cap C =  \{(0, 1, 1\}$$
So, $$\mathbb{P}(A \cap B) = \frac{1}{4} = \mathbb{P}(A) \cdot \mathbb{P}(B)$$
$$\mathbb{P}(A \cap C) = \frac{1}{4} = \mathbb{P}(A) \cdot \mathbb{P}(C)$$
$$\mathbb{P}(B \cap C) = \frac{1}{4} = \mathbb{P}(B) \cdot \mathbb{P}(C)$$

*(Q)* What is $A \cap B \cap C$? What is $\mathbb{P}(A \cap B \cap C)$

*(A)* Since, $$A \cap B \cap C = \phi$$
So, $$\mathbb{P}(A \cap B \cap C) = 0$$

*(Q)* Are the events $A$, $B$, $C$ mutually independent?

*(A)* Since,$$\mathbb{P}(A) \cdot \mathbb{P}(B) \cdot \mathbb{P}(C) = \frac{1}{2} \cdot \frac{1}{2} \cdot \frac{1}{2} = \frac{1}{8}$$
We have, $$\mathbb{P}(A \cap B \cap C) = 0 \neq \frac{1}{8} = \mathbb{P}(A) \cdot \mathbb{P}(B) \cdot \mathbb{P}(C)$$
Hence, these events cannot be mutually independent.


### 6 The Monty hall problem (*)
Consider the following game:

At a game show there are three seemingly identical doors. Behind one of the doors 
is a car, and behind the remaining two is a goat. The contestant of the game first 
gets to choose one of the three doors. The host then opens one of the other two 
doors to reveal a goat. The contestant now get’s a chance to either (a) switch 
their choice to the other unopened door or (b) stick to their original choice. 
The host then opens the door corresponding to the contestant’s final choice. They 
get to keep whatever is behind their final choice of door.

Does the contestant improve their chances of winning the car if they switch their 
choice? For clarity, we make the following assumptions:

1. The car is assigned to one of the doors at random with equal probability for 
each door.
2. The assignment of the car and the initial choice of the contestant are independent.
3. Once the contestant makes their initial choice, the host always opens a door 
which (a) has a goat behind it and (b) is not the contestant’s initial choice. 
If there is more than one such door (i.e. when the contestant’s initial choice 
corresponds to the door with a car behind it) the host chooses at random from the
two possibilities with equal probability.

To formalise our problem we introduce the following events for $i$ = 1, 2, 3:

- $A_i$ denotes event car is placed behind the $i$-th door;
- $B_i$ denotes event contestant initially chooses the $i$-th door;
- $C_i$ denotes event the host opens the $i$-th door to reveal a goat.

*(Q)* Consider a situation in which the contestant initially selects the first 
door ($B1$) and then the host opens the second door to reveal a goat ($C2$). What is 
$\mathbb{P}(A3 \mid B1 \cap C2)$?

*(A)*
We first compute $\mathbb{P}(A_3 \mid B_1 \cap C_2)$, which is:
$$
\mathbb{P}(A_3 \mid B_1 \cap C_2) = \frac{\mathbb{P}(C_2 \cap A_3 \cap B_1)}{\mathbb{P}(C_2 \cap B_1)}
= \frac{\mathbb{P}(C_2 \mid A_3 \cap B_1) \cdot \mathbb{P}(A_3 \cap B_1)}{\mathbb{P}(C_2 \cap B_1)}
= \frac{\mathbb{P}(C_2 \mid A_3 \cap B_1) \cdot \mathbb{P}(A_3 \mid B_1)}{\mathbb{P}(C_2 \mid B_1)}
$$
And we shall use the above formula.

To evaluate this expression we first need to compute $\mathbb{P}(C_2 \mid B_1)$:


$\mathbb{P}(C_2 \mid B_1) = \frac{\mathbb{P}(C_2 \cap B_1)}{\mathbb{P}(B_1)}$

$\qquad \qquad \; = \frac{\mathbb{P}(A_1 \cap C_2 \cap B_1) + \mathbb{P}(A_2 \cap C_2 \cap B_1) + \mathbb{P}(A_3 \cap C_2 \cap B_1)}{\mathbb{P}(B_1)}$

$\qquad \qquad \; = \frac{\mathbb{P}(A_1 \mid C_2 \cap B_1) \cdot \mathbb{P}(A_1 \cap B_1) + \mathbb{P}(A_2 \mid C_2 \cap B_1) \cdot \mathbb{P}(A_2 \cap B_1) + \mathbb{P}(A_3 \mid C_2 \cap B_1) \cdot \mathbb{P}(A_3 \cap B_1)}{\mathbb{P}(B_1)}$

$\qquad \qquad \; = \mathbb{P}(C_2 \mid A_1 \cap B_1) \cdot \mathbb{P}(A_1) + \mathbb{P}(C_2 \mid A_2 \cap B_1) \cdot \mathbb{P}(A_2) + \mathbb{P}(C_2 \mid A_3 \cap B_1) \cdot \mathbb{P}(A_3)$

$\qquad \qquad \; = \frac{1}{3} \cdot \{\mathbb{P}(C_2 \mid A_1 \cap B_1) + \mathbb{P}(C_2 \mid A_2 \cap B_1) + \mathbb{P}(C_2 \mid A_3 \cap B_1)\}$

$\qquad \qquad \; = \frac{1}{3} \cdot (\frac{1}{2} + 0 + 1) = \frac{1}{2}$

And we have $\mathbb{P}(C_2 \mid A_3 \cap B_1) = 1$, 

By using the independence of the choice of the door for the car and the contestnt's initial choice,
combined with the fact that the initial assignment of the car is random with equal probabilities,
we have $\mathbb{P}(A_3 \mid B_1) = \mathbb{P}(A_3) = \frac{1}{3}$.

Hence, we have
$$
\mathbb{P}(A_3 \mid B_1 \cap C_2) = = \frac{\mathbb{P}(C_2 \mid A_3 \cap B_1) \cdot \mathbb{P}(A_3 \mid B_1)}{\mathbb{P}(C_2 \mid B_1)} = \frac{1 \cdot \frac{1}{3}}{\frac{1}{2}} = \frac{2}{3}
$$

*(Q)* What does this suggest about a good strategy? Should we switch choices?

*(A)* By the same argument we obtain $\mathbb{P}(A_{i_3} \mid B_{i_1} \cap C_{i_2}) = \frac{2}{3}$
for any $i_1, i_2, i_3 \in \{1,2,3\}$ such that $\{i_1,i_2,i_3\} = \{1,2,3\}$. 
Hence, the contestant always increases their chances by switching choice (from $i_1$ to $i_3$).


### 7 A game of marbles (**)

This question is left as an optional challenge for those who have completed all 
remaining questions.

Suppose there two players - $A$ and $B$. Player $A$ has $n$ marbles and player $B$ 
has $k − n$ marbles $(n \leq k)$, so the two players have $k$ marbles between them. 
The game proceeds through a series of rounds. In each round $A$ has probability $p$ 
of winning the round, and $B$ has probability $1 − p$ of winning the round, where 
$p \in (0, 1/2)$ is a fixed number. If $A$ wins a round then $B$ must give $A$ 
one of his marbles. If $B$ wins the round then $A$ must give $B$ one of his marbles. 
All of the rounds are independent from one another. A player wins the entire 
game of marbles when they have all $k$ marbles, and the other player has none.

*(Q)* Can you give a formula for the probability that player $A$ wins the game 
in terms of $n$ and $k$ and $p$?




