---
title: "myFirstRMarkdown"
author: "Kexin Wu"
date: "2021/10/4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Wave plot

Define a vector called x consisting of a sequence which starts at 0 and goes to 20 in increments of 0.01.
```{r}
x <- seq(0, 20, 0.01)
```

Next create a vector called y which is of the same length of x and is such that the ith entry of y is equal to the sin function of the ith entry of x.
```{r}
y <- sin(x)
```

Now create a dataframe called “sin_df” with two coloumns: x and y. You can inspect the first few rows of your dataframe with the head() function like this:
head(sin_df,3)
```{r}
sin_df <- data.frame(x, y)
head(sin_df, 3)
```

## Including plots
```{r}

y <- sin(x)
plot(x, y, type = "b", asp = 4)
```

## some formulae


$y=\sin(x)$


\[ \sin(x)=\sum_{n=1}^{\infty}(-1)^{n+1}\cdot \frac{x^{2n-1}}{(2n-1)!}
\approx x-\frac{x^3}{3!}+\frac{x^5}{5!}-\frac{x^7}{7!}\ldots. \]


\[ A=\left(\begin{matrix} a_{11} & a_{12}\\ a_{21} & a_{22} \end{matrix}\right). \]