---
title: "Notes of Lecture6"
author: "Kexin Wu"
date: "10/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lecture 6: Tidy data and iteration

# What is tidy data

Two features of tidy data:

1. Each row corresponds to a specific and unique observation representing a similar sort of thing

2. Columns correspond to single vatiables with the same sort of value for each observaion 

# Reshaping data
 Narrow data to wide data
 
 We can use pivot functions to efficiently reshape our data
 
pivot_wider(names_from = , values_from = )

pivot_longer(cols = c(), names_to = , values_to = )


Example: How can we compute the win rate per participant?
```{r}
library(tidyverse)
name = c("Alice", "Bob", "Charlie")
wins_of_2018 = c(5, 9, 3)
wins_of_2019 = c(9, 7, 2)
wins_df_wide <- data.frame(name, wins_of_2018, wins_of_2019) %>%
  rename('2018' = wins_of_2018, '2019' = wins_of_2019)
losses_of_2018 = c(7, 15, 12)
losses_of_2019 = c(8, 4, 10)
losses_df_wide <- data.frame(name, losses_of_2018, losses_of_2019) %>%
  rename('2018' = losses_of_2018, '2019' = losses_of_2019)

wins_df_wide
wins_df_narrow <- wins_df_wide %>%
   pivot_longer(!name, names_to = "year", values_to = "wins")
wins_df_narrow

losses_df_wide
losses_df_narrow <- losses_df_wide %>%
  pivot_longer(!name, names_to = 'year', values_to = 'losses')
losses_df_narrow

wins_losses_df <- inner_join(wins_df_narrow, losses_df_narrow) %>%
  mutate(win_rate = wins / (wins + losses))
wins_losses_df
```

# Uniting and seperating data
 Multiple variables within a column
```{r}
wins_over_total_df <- inner_join(wins_df_narrow, losses_df_narrow) %>%
  mutate(w_over_t = paste(wins, "/", (wins + losses))) %>%
  select(name, year, w_over_t)
wins_over_total_df

sep_df <- wins_over_total_df %>%
  separate(w_over_t, into = c("wins", "totals"), sep = "/")
sep_df
# Note, "wins" and "totals" are not numerical variables 
# Use the convert to convert those string variables into numerical variables
sep_df <- wins_over_total_df %>%
  separate(w_over_t, into = c("wins", "totals"), sep = "/", convert = TRUE)
sep_df

sep_df %>%
  mutate(losses = totals - wins) %>%
  select(-totals)

# Unite function
uni_df <- sep_df %>%
   unite(w_over_t, wins, totals, sep = "/")
uni_df
```

# Iteration

 The standard approach to iteration is through loops e.g. for, while etc.
 
 In R we prefer vectorized and map operations where possible:
 
 1. Historically, this led to efficiency improvements. However, this is often no longer the case.
 
 2. Map based approaches are typically more readable. 
 
# The map function
```{r}
is_div_2_3 <- function(x) {
  if (x %% 2 == 0){
    return(TRUE)
  } else if(x %% 3 == 0){
    return(TRUE)
  } else{
    return(FALSE)
  }
}
is_div_2_3(3)
v <- c(1, 2, 3, 4, 5)
is_div_2_3(v)
map(v, is_div_2_3)
```
 The output of map() is a list of function values.

 The output of map_lgl() is a vector of Booleans.
```{r}
map_lgl(v, is_div_2_3)
```
 The output of map_dbl() is a vector of doubles.
```{r}
example_f <- function(x) {
  if(is_div_2_3(x)){
    return(x)
  } else{
    return(0)
  }
}
v <- c(1, 2, 3, 4, 5)
map_dbl(v, example_f)
```
 The output of map_chr() is a vector of strings.
```{r}
library(english)
example_f_eng <- function(x){
  as.character(as.english(example_f(x)))
}
v <- c(1, 2, 3, 4, 5)
map_chr(v, example_f_eng)
```

# Vectorization
```{r}
is_div_2_3_vect <- function(x){
 return(x %% 2 == 0 | x %% 3 == 0)
}
is_div_2_3_vect(v)
```
```{r}
example_f_vect <- function(x){
  return(x * is_div_2_3_vect(x))
}
example_f_vect(v)
```
```{r} 
example_f_eng_vect <- function(x){
  return(as.english(example_f_vect(v)))
}
example_f_eng_vect(v)
```

# Nesting and unnesting
```{r}
musicians <- full_join(band_members, band_instruments)
musicians
musicians_nasted <- musicians %>% 
  group_by(name) %>% 
  nest()
musicians_nasted
musicians_nasted %>% 
  unnest(cols = data)
```

 Our goal is to create a function which:
 
 1) Takes as input a data frame and a variable name
 
 2) Computes the correlation with all other numeric variables
 
 3) Returns the name of the variable with maximal absolute correlation, and the corresponding correlation
 
```{r}
max_cor_var <- function(df, col_name){ # function to determine the variable with maximal correlation
  v_col <- -df %>% select(all_of(col_name)) # extract variable based on col_name
  
  df_num <- -df %>%
    select_if(is.numeric) %>%
    select(-all_of(col_name)) # select all numeric variables excluding col_name
  
  correlations <- unlist(map(df_num, function(x){
    cor(x, v_col, use = "complete.obs")
  })) # compute correlations with all other numeric variable
  
  max_abs_cor_var <- names(which(abs(correlations) == max(abs(correlations)))) # extract the variable name
  cor <- as.double(correlations[max_abs_cor_var]) # compute the correlation
  
  return(data.frame(var_name = max_abs_cor_var, cor = cor)) # return dataframe
}

library(palmerpenguins)
penguins %>%
  max_cor_var("body_mass_g")

penguins %>%
  group_by(species) %>%
  nest() %>%
  mutate(max_cor = map(data, ~max_cor_var(.x, "body_mass_g"))) %>%
  select(-data) %>%
  unnest(cols = max_cor)
``` 

# Missing data
 Missing data is remarkably common in practical Data Science applications.
 
 1. The value of an indiviudal variable replaced with "NA" (not available).
  
 2. Implicit missing data: Entire rows missing from a data frame.
 
# Making missing data explicit

complete.cases()

filter(complete.cases(.))

# Imputation by mean
```{r}
impute_by_mean <- function(x){
  mu <- mean(x, na.rm = 1) # first compute the mean of x
  impute_f <- function(z){ # coordinate_wise imputation
    if(is.na(z)){
      return(mu) # if z is na replace with mean
    }else{
      return(z) # otherwise leave in place
    }
  }
  return(map_dbl(x, impute_f)) # apply the map function to impute across vector
}
x <- c(1, 2, NA, 4)
impute_by_mean(x)
```